Simulating an Isolated Dwarf Galaxy with Energetic Feedback and Chemical Yields from Individual Stars

### 2018-07-15 main.tex
wording changes to abstract and conclusion

\begin{abstract}
In order to better understand the relationship between feedback and galactic chemical evolution, we have developed a new model for stellar feedback at grid resolutions of only a few parsecs in global disk simulations, using the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. For the first time in galaxy scale simulations, we simulate detailed stellar feedback from individual stars including asymptotic giant branch winds, photoelectric heating, Lyman-Werner dissociating radiation, ionizing radiation tracked through an adaptive ray-tracing radiative transfer method, and core collapse and Type Ia supernovae. We furthermore follow the star-by-star chemical yields using tracer fields for 15 metal species: C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba. We include the yields ejected in massive stellar winds, but greatly reduce the winds' velocities due to computational constraints. We describe these methods in detail in this work and present the first results from 500~Myr of evolution of an isolated dwarf galaxy with properties similar to a Local Group ultrafaint dwarf galaxy. We demonstrate that our physics and feedback model is capable of producing a dwarf galaxy whose evolution is consistent with observations in both the Kennicutt-Schmidt relationship and extended Schmidt relationship. Effective feedback \textbf{drives outflows with a greater metallicity than the ISM}, leading to low metal retention fractions consistent with observations. Finally, we demonstrate that these simulations yield valuable information on the variation in mixing behavior of individual metal species within the multi-phase interstellar medium.
\end{abstract}

@@

\subsubsection{Differential Evolution of Elements Within the ISM}
It is important to understand how metals from each source of stellar yields enrich the ISM. Observations of more massive dwarf galaxies than those simulated here indicate fairly uniform radial gas-phase metallicity profiles, even beyond the stellar radius \citep[e.g.][]{Werk2011,Belfiore2017}. This requires that metal mixing and transport occur on hundred megayear timescales, much more rapidly than the gigayear timescale expected from assuming transport at the cold gas sound speed. Therefore, either metals are transported first through a hot phase with high sound speed, or through efficient turbulent mixing within the ISM \citep[e.g.][]{Avillez2002,Tassis2008,YangKrumholz2012}. It remains uncertain how metal abundances vary in detail within these galaxies, beyond one-dimensional radial profiles, and whether or not abundance distributions depend on the metal species. It is even more unclear how metals are transported and distributed within ultrafaint dwarf galaxies, which generally host too few \ion{H}{2} regions for a detailed examination.

@@

\section{Conclusion}
\label{sec:conclusion}
We have developed a new method for simulating galaxy evolution with detailed feedback and chemical enrichment. For the first time on galaxy scales, we simultaneously model multi-channel stellar feedback in detail, using individual star particles to model core collapse and Type Ia SNe, ionizing radiation followed through radiative transfer, photoelectric heating, Lyman-Werner radiation and pollution from AGB and massive stellar winds. This treatment of feedback, coupled with the detailed chemistry and heating/cooling physics followed with \textsc{Grackle}, allows us to capture realistic galaxy evolution in detail. In this work, we apply these methods to simulate the evolution of an isolated, low-mass, dwarf galaxy modeled after the $z=0$ properties of the Leo P ultrafaint dwarf galaxy. We present an overview of the properties of this simulation in this work.

For our simulated dwarf galaxy, we find:
\begin{enumerate}
\item Multi-channel feedback is effective in this galaxy in regulating star formation to occur at a rate consistent with the Kennicutt-Schmidt relationship and the extended Schmidt law in observed galaxies. (See Figures~\ref{fig:sfr_mass_evolution} and \ref{fig:KS}).

@@

\item 
Only $\sim$4\% of metals are retained in the disk of our simulated galaxy, consistent with the observed metal retention fractions of ultrafaint dwarfs.  By the end of the simulation $\sim$45\% of the remaining metals remain within the virial radius (but outside the galaxy) and $\sim$50\% 
have been ejected beyond the virial radius. (See Figures \ref{fig:mass_outflow}, \ref{fig:outflow_velocity}, and \ref{fig:metal_evolution}.)

\item 
The gas scale height is thin ($\sim 50$~pc), yet resolved, 
beyond our stellar radius, with larger scale heights ($\sim 100$--200~pc) driven by feedback interior to the stellar radius. This is comparable to the resolution limit of the diffuse HI in observed, gaseous ultrafaint dwarfs. At a spatial resolution of 100~pc, our galaxy has a peak HI column density $N_{\rm HI} = 2.8-4.3 \times 10^{20}$~cm$^{-2}$, depending on inclination (See Figure~\ref{fig:scale_height}).

\item The ISRF of our galaxy varies strongly in both space and time by orders of magnitude. It is unclear how important these fluctuations are as a source of feedback, or if the affect can be approximated with a time-averaged radial profile. (See Figures \ref{fig:ISRF} and \ref{fig:ISRF_2D}.)

@@

\item Finally, we used O and N as examples to briefly demonstrate that there are marked differences in how individual metal species are distributed within the ISM of our galaxy. These variations could be tied to differences in elemental yields among different sources (for example, AGB winds vs. SNe), as suggested by \cite{KrumholzTing2018}. This will be explored in more detail in future work (Emerick et. al. \textit{in prep.}).
\end{enumerate}

>>>>

\begin{abstract}
In order to better understand the relationship between feedback and galactic chemical evolution, we have developed a new model for stellar feedback at grid resolutions of only a few parsecs in global disk simulations, using the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. For the first time in galaxy scale simulations, we simulate detailed stellar feedback from individual stars including asymptotic giant branch winds, photoelectric heating, Lyman-Werner radiation, ionizing radiation tracked through an adaptive ray-tracing radiative transfer method, and core collapse and Type Ia supernovae. We furthermore follow the star-by-star chemical yields using tracer fields for 15 metal species: C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba. We include the yields ejected in massive stellar winds, but greatly reduce the winds' velocities due to computational constraints. We describe these methods in detail in this work and present the first results from 500~Myr of evolution of an isolated dwarf galaxy with properties similar to a Local Group, low-mass dwarf galaxy. We demonstrate that our physics and feedback model is capable of producing a dwarf galaxy whose evolution is consistent with observations in both the Kennicutt-Schmidt relationship and extended Schmidt relationship. Effective feedback \textbf{drives outflows with a greater metallicity than the ISM}, leading to low metal retention fractions consistent with observations. Finally, we demonstrate that these simulations yield valuable information on the variation in mixing behavior of individual metal species within the multi-phase interstellar medium.
\end{abstract}

@@

\subsubsection{Differential Evolution of Elements Within the ISM}
It is important to understand how metals from each source of stellar yields enrich the ISM. Observations of more massive dwarf galaxies than those simulated here indicate fairly uniform radial gas-phase metallicity profiles, even beyond the stellar radius \citep[e.g.][]{Werk2011,Belfiore2017}. This requires that metal mixing and transport occur on hundred megayear timescales, much more rapidly than the gigayear timescale expected from assuming transport at the cold gas sound speed. Therefore, either metals are transported first through a hot phase with high sound speed, or through efficient turbulent mixing within the ISM \citep[e.g.][]{Avillez2002,Tassis2008,YangKrumholz2012}. It remains uncertain how metal abundances vary in detail within these galaxies, beyond one-dimensional radial profiles, and whether or not abundance distributions depend on the metal species. It is even more unclear how metals are transported and distributed within low-mass dwarf galaxies, which generally host too few \ion{H}{2} regions for a detailed examination.

@@

\section{Conclusion}
\label{sec:conclusion}
We have developed a new method for simulating galaxy evolution with detailed feedback and chemical enrichment. For the first time on galaxy scales, we simultaneously model multi-channel stellar feedback in detail, using individual star particles to model core collapse and Type Ia SNe, ionizing radiation followed through radiative transfer, photoelectric heating, Lyman-Werner radiation and pollution from AGB and massive stellar winds. This treatment of feedback, coupled with the detailed chemistry and heating/cooling physics followed with \textsc{Grackle}, allows us to capture realistic galaxy evolution in detail. In this work, we apply these methods to simulate the evolution of an isolated, low-mass, dwarf galaxy modeled after the $z=0$ properties of the Leo P dwarf galaxy. We present an overview of the properties of this simulation in this work.

For our simulated dwarf galaxy, we find:
\begin{enumerate}
\item Multi-channel feedback is effective in regulating star formation to a rate consistent with the Kennicutt-Schmidt relationship and the extended Schmidt law in observed galaxies. (See Figures~\ref{fig:sfr_mass_evolution} and \ref{fig:KS}).

@@

\item 
Only $\sim$4\% of metals are retained in the disk of our simulated galaxy, consistent with the observed metal retention fractions of low-mass dwarfs.  By the end of the simulation $\sim$45\% of the remaining metals stay within the virial radius (but outside the galaxy), while $\sim$50\% 
have been ejected beyond the virial radius. (See Figures \ref{fig:mass_outflow}, \ref{fig:outflow_velocity}, and \ref{fig:metal_evolution}.)

\item 
Beyond the stellar radius, the gas scale height is thin ($\sim 50$~pc), yet resolved, with larger scale heights ($\sim 100$--200~pc) driven by feedback interior to the stellar radius. This is comparable to the resolution limit of the diffuse HI in observed, gaseous low-mass dwarfs. At a spatial resolution of 100~pc, our galaxy has a peak HI column density $N_{\rm HI} = 2.8-4.3 \times 10^{20}$~cm$^{-2}$, depending on inclination (See Figure~\ref{fig:scale_height}).

\item The ISRF of our galaxy varies strongly in both space and time by orders of magnitude. It is unclear how important these fluctuations are as a source of feedback, or if the affect can be approximated with a time-averaged radial profile. The importance of radiation feedback in our model will be investigated in a future work. (See Figures \ref{fig:ISRF} and \ref{fig:ISRF_2D}.)

@@

\item Finally, we present gas-phase oxygen and nitrogen distributions as examples to briefly demonstrate that there are marked differences in how individual metal species are distributed within the ISM of our galaxy. These variations could be tied to differences in elemental yields among different sources (for example, AGB winds vs. SNe), as suggested by \cite{KrumholzTing2018}. This will be explored in more detail in future work (Emerick et. al. \textit{in prep.}).
\end{enumerate}
